
Initialising the required libraries

```{r}
# Library Initialisation
library(DMwR)
library(data.table)
library(Information)
library(gridExtra)
library(class)
library(caret)
library(MASS)

```

Reading the data set

```{r}
# Reading the dataset
application_test <- read.csv("C:/University of Chicago/Data Mining/Project/application_test.csv", header=TRUE,stringsAsFactors=FALSE)
application_train <- read.csv("C:/University of Chicago/Data Mining/Project/application_train.csv", header=TRUE,stringsAsFactors=FALSE)

```


Imputing the missing/NA values

```{r}
# Categorizing the OCCUPATION_TYPE column in application_train
application_train$OCCUPATION_TYPE <- as.integer(as.numeric(as.factor(application_train$OCCUPATION_TYPE)))
# Categorizing the ORGANIZATION_TYPE COLUMN in application_train
application_train$ORGANIZATION_TYPE <- as.integer(as.numeric(as.factor(application_train$ORGANIZATION_TYPE)))
# Categorizing the NAME_CONTRACT_TYPE column in application_train
application_train$NAME_CONTRACT_TYPE <- as.integer(as.numeric(as.factor(application_train$NAME_CONTRACT_TYPE)))
# Categorizing the CODE_GENDER column in application_train
application_train$CODE_GENDER <- as.integer(as.numeric(as.factor(application_train$CODE_GENDER)))
# Categorizing the EMERGENCYSTATE_MODE column in application_train
application_train$EMERGENCYSTATE_MODE <- as.integer(as.numeric(as.factor(application_train$EMERGENCYSTATE_MODE)))
# Categorizing the EMERGENCYSTATE_MODE column in application_train
application_train$EMERGENCYSTATE_MODE <- as.integer(as.numeric(as.factor(application_train$EMERGENCYSTATE_MODE)))
# Categorizing the FLAG_OWN_CAR  column in application_train
application_train$FLAG_OWN_CAR  <- as.integer(as.numeric(as.factor(application_train$FLAG_OWN_CAR)))
# Categorizing the FLAG_OWN_REALTY  column in application_train
application_train$FLAG_OWN_REALTY  <- as.integer(as.numeric(as.factor(application_train$FLAG_OWN_REALTY)))
# Categorizing the NAME_TYPE_SUITE  column in application_train
application_train$NAME_TYPE_SUITE  <- as.integer(as.numeric(as.factor(application_train$NAME_TYPE_SUITE)))
# Categorizing the NAME_INCOME_TYPE  column in application_train
application_train$NAME_INCOME_TYPE  <- as.integer(as.numeric(as.factor(application_train$NAME_INCOME_TYPE)))
# Categorizing the NAME_EDUCATION_TYPE    column in application_train
application_train$NAME_EDUCATION_TYPE   <- as.integer(as.numeric(as.factor(application_train$NAME_EDUCATION_TYPE)))
# Categorizing the NAME_FAMILY_STATUS    column in application_train
application_train$NAME_FAMILY_STATUS   <- as.integer(as.numeric(as.factor(application_train$NAME_FAMILY_STATUS)))
# Categorizing the NAME_HOUSING_TYPE    column in application_train
application_train$NAME_HOUSING_TYPE   <- as.integer(as.numeric(as.factor(application_train$NAME_HOUSING_TYPE)))
# Categorizing the WEEKDAY_APPR_PROCESS_START    column in application_train
application_train$WEEKDAY_APPR_PROCESS_START   <- as.integer(as.numeric(as.factor(application_train$WEEKDAY_APPR_PROCESS_START)))
# Converting the NULL Values to NA
application_train$HOUSETYPE_MODE <- ifelse(application_train$HOUSETYPE_MODE=="","NA",application_train$HOUSETYPE_MODE)
# Converting the NULL Values to NA
application_train$WALLSMATERIAL_MODE<- ifelse(application_train$WALLSMATERIAL_MODE=="","NA",application_train$WALLSMATERIAL_MODE)
# Converting the NULL Values to NA
application_train$EMERGENCYSTATE_MODE<- ifelse(application_train$EMERGENCYSTATE_MODE=="","NA",application_train$EMERGENCYSTATE_MODE)
# Subsetting the Numeric class columns
subset_colclasses <- function(DF, colclasses="numeric") {
  DF[,sapply(DF, function(vec, test) class(vec) %in% test, test=colclasses)]
}
# Subsetting the data into numeric, integer, character
subset_numeric <- subset_colclasses(application_train,colclasses="numeric")
subset_integer <- subset_colclasses(application_train,colclasses="integer")
subset_character <- subset_colclasses(application_train,colclasses="character")
f=function(x){
   x<-as.numeric(as.character(x)) #first convert each column into numeric if it is from factor
   x[is.na(x)] =mean(x, na.rm=TRUE) #convert the item with NA to mean value from the column
   x #display the column
}
ss=data.frame(apply(subset_numeric,2,f))
f_median=function(x){
   x<-as.integer(as.character(x)) #first convert each column into numeric if it is from factor
   x[is.na(x)] =median(x, na.rm=TRUE) #convert the item with NA to median value from the column
   x #display the column
}
ss_int=data.frame(apply(subset_integer,2,f))
uniqueX <- unique(subset_character$FONDKAPREMONT_MODE)
uniqueX <- uniqueX[uniqueX!="NA"]
subset_character$FONDKAPREMONT_MODE[subset_character$FONDKAPREMONT_MODE=="NA"] = sample(uniqueX,1) #convert the item with NA to median value from the column
uniqueY <- unique(subset_character$HOUSETYPE_MODE)
uniqueY <- uniqueY[uniqueY!="NA"]
subset_character$HOUSETYPE_MODE[subset_character$HOUSETYPE_MODE=="NA"] = sample(uniqueY,1) #convert the item with NA to median value from the column
uniqueZ <- unique(subset_character$WALLSMATERIAL_MODE)
uniqueZ <- uniqueZ[uniqueZ!="NA"]
subset_character$WALLSMATERIAL_MODE[subset_character$WALLSMATERIAL_MODE=="NA"] = sample(uniqueZ,1) #convert the item with NA to median value from the column
subset_master_character <- apply(subset_character,2,function(x)as.integer(as.numeric(as.factor(x))))
subset_master_character <- data.frame(subset_master_character)

```

Creating the cleaned Master data file

```{r}
# Master Data
imputed_master <- cbind(ss,ss_int,subset_master_character)
#write.csv(imputed_master,"C:/University of Chicago/Data Mining/Project/TrainDataCleaned.csv")

```

Scaling the cleaned data and applying PCA on it

```{r}
#TrainDataCleaned = read.csv("C:/University of Chicago/Data Mining/Project/TrainDataCleaned.csv")
TrainDataCleaned = imputed_master
TrainDataCleaned =TrainDataCleaned[,-1]
#str(TrainDataCleaned)
#names(TrainDataCleaned)
Target= TrainDataCleaned[,67]
TrainDataCleaned = TrainDataCleaned[,-c(66,67)]
#str(TrainDataCleaned)
TrainDataCleaned <- scale(TrainDataCleaned,center=apply(TrainDataCleaned,2,mean),scale=apply(TrainDataCleaned,2,sd))


Project.Data.PCA <- princomp(TrainDataCleaned)



barplot(Project.Data.PCA$sdev^2/sum(Project.Data.PCA$sdev^2), ylim=c(0,1))


cumsum(Project.Data.PCA$sdev^2/sum(Project.Data.PCA$sdev^2))
plot(cumsum(Project.Data.PCA$sdev^2/sum(Project.Data.PCA$sdev^2)))

```


Selecting the first 65 components to get about 90% of the total variance, and setting the target as the first variable

```{r}
TrainDataReducedPCA <- Project.Data.PCA$scores[,1:65] 

TrainDataReducedPCA <- cbind.data.frame(Target,TrainDataReducedPCA)
#str(TrainDataReducedPCA)

```

Splitting this data into train and test and converting the Target into factors

```{r}
set.seed(3)
n=round(length(TrainDataReducedPCA[,1])*0.7)
totaln=length(TrainDataReducedPCA[,1])

totaln
n

all =1:n
train = sample(1:totaln,size =n,replace = FALSE)
train =sort(train)
hold= all[-train]


TrainPCA <- as.data.frame(TrainDataReducedPCA[train,])
TestPCA <- as.data.frame(TrainDataReducedPCA[hold,])
#str(TrainPCA)

TrainPCA$Target =factor(TrainPCA$Target)
TestPCA$Target <- factor(TestPCA$Target)

```


Applying SMOTE on the train and test samples to correct the imbalance in the response

```{r}
TrainDataSmotePCA <- SMOTE(Target~., data=TrainPCA,perc.over = 200,k=5,learner=NULL) 
TestDataSmotePCA <-  SMOTE(Target~., data=TestPCA,perc.over = 200,k=5,learner=NULL)

#write.csv(TrainDataSmotePCA,"C:/University of Chicago/Data Mining/Project/TrainDataSmotePCA.csv")
#write.csv(TestDataSmotePCA,"C:/University of Chicago/Data Mining/Project/TestDataSmotePCA.csv")

```


Applying LDA to the train and test post SMOTE

```{r}
ldatrain <- lda(Target~., data=TrainDataSmotePCA)
ldatrainclass<- predict(ldatrain)$class

confusionMatrix(ldatrainclass,TrainDataSmotePCA$Target)


ldatestclass <- predict(ldatrain, newdata=TestDataSmotePCA[,-1])$class
confusionMatrix(ldatestclass,TestDataSmotePCA$Target)



```


Analysing the group means of the two LDA classes

```{r}
ldatrainGroupMeans <- as.data.frame(t(ldatrain$means))


names(ldatrainGroupMeans)<- c("Group0","Group1")

ldatrainGroupMeansby0 <- ldatrainGroupMeans[order(-ldatrainGroupMeans[,1]),] 
ldatrainGroupMeansby1 <- ldatrainGroupMeans[order(-ldatrainGroupMeans[,2]),] 

head(ldatrainGroupMeansby0,10)
head(ldatrainGroupMeansby1,10)

```


```{r}
#Significant variables for predicting 0
#Project.Data.PCA$loadings
PCALoadings <- as.data.frame(Project.Data.PCA$loadings[,119])

#str(Project.Data.PCA$loadings)
library(data.table)
library(spatial.tools)
setDT(PCALoadings, keep.rownames = TRUE)[]

#biplot(Project.Data.PCA$scores[1:10,c(1,4)],Project.Data.PCA$loadings[,c(1,4)])

```

```{r}
apply(Project.Data.PCA$loadings[,c(3,4,25,1,20)],2, function(x) PCALoadings[which.max(x),1])

#apply(Project.Data.PCA$loadings[,c(8,13,16)],2, function(x) PCALoadings[which.max(x),1])


```

```{r}
apply(Project.Data.PCA$loadings[,c(2,19,15,14,24)],2, function(x) PCALoadings[which.max(x),1])

#apply(Project.Data.PCA$loadings[,c(24,31,46)],2, function(x) PCALoadings[which.max(x),1])


```

Applying KNN on the Train and Test

```{r}
#str(TrainDataSmotePCA)
#str(TestDataSmotePCA)
knnoutput2 <- list()
for(i in 1:10) {
  knnoutput2[[i]] =knn(TrainDataSmotePCA[,-1], TestDataSmotePCA[,-1],TrainDataSmotePCA[,1], k=i)
}


```

Extracting the accuracy, kappa, sensitivity and specificity from the knn results

```{r}
sensitivity =c()
specificity=c()
accuracy =c()
kappa =c()
for(i in 1:10) {
  sensitivity[i]=confusionMatrix(knnoutput2[[i]],TestDataSmotePCA$Target)$byClass[[1]]
  specificity[i]=confusionMatrix(knnoutput2[[i]],TestDataSmotePCA$Target)$byClass[[2]]
  accuracy[i]=confusionMatrix(knnoutput2[[i]],TestDataSmotePCA$Target)$overall[[1]]
  kappa[i]=confusionMatrix(knnoutput2[[i]],TestDataSmotePCA$Target)$overall[[2]]
}

```

Plotting the accuracy vs. number of neighbours for KNN

```{r}
accuracy
sensitivity
specificity
kappa

plot(1:10,accuracy, type="l", xlab="number of neighbours", ylab="Accuracy", main="KNN Accuracy vs. number of neighbours")

```

